Language-Based Security for Low-Level MPC

### 2024-05-17 ppdp24/intro.tex
checkpoint

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. The security semantics of
MPC include both confidentiality and integrity properties incorporated
into models such as real/ideal (aka simulator) security and universal
composability (UC), developed primarily by the cryptography community
\cite{XXX}.  Related proof methods are well-studied \cite{Lindell2017}
but mostly manual. Somewhat independently, a significant
body of work in programming languages has focused on definition and
enforcement of confidentiality and integrity \emph{hyperproperties}
\cite{10.5555/1891823.1891830} such as noninterference and delimited
release \cite{XXX}. Following a tradition of connecting
cryptographic and PL-based security models \cite{XXX}, recent work has
also recognized connections between MPC security models and
hyperpoperties of, e.g., noninterference \cite{8429300}, and even
leveraged these connections to enforce MPC security through mechanisms
such as security types \cite{XXX}. Major benefits of this connection
in an MPC setting include better language abstractions for defining
protocols and for mechanisation and even automation of security proofs.
The goal of this paper is to develop a PL model for defining a variety of
low-level probabilistic MPC protocols, to formulate a collection
of confidentiality and integrity hyperproperties for our model
model with familiar information flow analogs, and to show how these
properties can be leveraged for improved proof automation.

The distinction between high- and low-level languages for MPC is
important. High-level languages such as Wysteria \cite{XXX} and
Viaduct \cite{XXX} are designed to provide effective programming of
full applications. These language designs incorporate sophisticated
verified compilation techniques such as orchestration \cite{XXX} to
guarantee high-level security properties, and they rely on
\emph{libraries} of low-level MPC protocols, such as binary and
arithmetic circuits. These low-level protocols encapsulate
abstractions such as secret sharing and semi-homomorphic encrypytion,
and rely on probabilistic progamming methods. So, low-level MPC
programming and protocol verification is a distinct challenge and both
critical to the general challenge of PL design for MPC and
complementary to high-level language design.

The connection between information flow hyperproperties and
MPC security is also complicated especially at a low level.
MPC protocols involve communication between a group of distributed
clients called a \emph{federation} that collaboratively compute
and publish the result of some known \emph{ideal functionality}
$\idealf$, maintaining confidentiality of inputs to $\idealf$ without
the use of a trusted third party. However, since the outputs of
$\idealf$ are public, some information about inputs is inevitably
leaked. Thus, the ideal
functionality establishes a declassification policy \cite{XXX}, which
is more difficult to enforce than pure noninterference.  And
subtleties of, e.g., semi-homomorphic encryption are central to both
confidentiality and integrity properties of protocols and similarly
difficult to track with coarse-grained security types alone.

Nevertheless, as previous authors have observed \cite{XXX}, low-level
protocol design patterns such as secret sharing and circuit gate
structure have compositional properties that can be independently
verified and then leveraged in larger proof contexts. We
contribute to this line of work by developing an automated
verification technique for subprotocols and show how it can be
integrated as a tactic in a larger security proof.

@@

As mentioned above, several high-level languages have been developed
for writing MPC applications, 
and frequently exploit the connection between hyperproperties and MPC
security. Previous work on analysis for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
properties of complex MPC circuits, in particular a user-friendly
specification and automated enforcement of declassification bounds in
programs that use MPC in subprograms. The Wys$^\star$ language
\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
similar goals and includes a trace-based semantics for reasoning about
the interactions of MPC protocols. Their compiler also guarantees that
underlying multi-threaded protocols enforce the single-threaded source
language semantics. These two lines of work were focused on passive
security. The Viaduct language \cite{XXX} has a well-developed
information flow type system that automatically enforces both
confidentiality and integrity through hyperproperties such as robust
declassification, in addition to rigorous compilation guarantees
through orchestration \cite{XXX}. However, these high level
languages lack probabilistic features and other abstractions of
low-level protocols, the implementation and security of which
are typically assumed as a selection of library compnonents.

Various related low-level languages with probabilistic features have
also been developed. The $\lambda_{\mathrm{obliv}}$ language
\cite{darais2019language} uses a type system is used to automatically
enforce so-called probabilistic trace obliviousness.  But similar to
previous work on oblivious data structures \cite{10.1145/3498713},
obliviousness is related to pure noninterference, not the relaxed form
related to passive MPC security. The Haskell-based security type system in
\cite{XXX} enforces a version of conditioned noninterence that is
sound for passive security in a probabilisitic low-level setting, but
they do not consider malicious security. And properties of real/ideal
passive and malicious security for a probabilistic language have been
formulated in EasyCrypt \cite{8429300}-- though their proof methods,
while mechanized, are fully manual, and their formulation of malicious
security is not as clearly related to robust declassification as the
one we present in Section \ref{section-hyperprop}.

Program logics for probabilistic languages and specifically reasoning
about properties such as joint probabilistic independence is also
important related work. Probabilistic Separation Logic (PSL)
\cite{barthe2019probabilistic} develops a logical framework for
reasoning about probabilistic independence (aka separation) in
programs, and they consider several (hyper)properties, such as perfect
secrecy of one-time-pads and indistinguishability in secret sharing,
that are critical to MPC. However, their methods are manual, and
don't include conditional independence (separation). This
latter issue has been addressed in Lilac \cite{XXX}. The
application of Lilac-style reasoning to MPC protocols has not
previously been explored, as we do in Section
\ref{section-example-gmw}.

@@

\paragraph{Fully and partially automated verification.} In Section
\ref{section-bruteforce} we develop a method for automatically
computing the pmf of $\minifed$ protocols in $\mathbb{F}_2$, that can
be automatically queried to enforce hyperperproperties of
security. This method is perfectly accurate but has high complexity;
we show this can be partially mitigated by conversion of protocols in
$\mathbb{F}_2$ to stratified Datalog which is amenable to HPC
acceleration. Furthermore, in Section \ref{section-examples-gmw} we
consider in detail how this automated technique can be used as a local
automated tactic for proving security in arbitrarily large GMW
circuits using conditional probabilistic independence as in
\cite{XXX}.

>>>>

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. The security semantics of
MPC include both confidentiality and integrity properties incorporated
into models such as real/ideal (aka simulator) security and universal
composability (UC), developed primarily by the cryptography community
\cite{evans2018pragmatic}.  Related proof methods are well-studied \cite{Lindell2017}
but mostly manual. Somewhat independently, a significant
body of work in programming languages has focused on definition and
enforcement of confidentiality and integrity \emph{hyperproperties}
\cite{10.5555/1891823.1891830} such as noninterference and delimited
release \cite{sabelfeld2009declassification}. Following a tradition of connecting
cryptographic and PL-based security models \cite{10.1007/3-540-44929-9_1}, recent work has
also recognized connections between MPC security models and
hyperpoperties of, e.g., noninterference \cite{8429300}, and even
leveraged these connections to enforce MPC security through mechanisms
such as security types \cite{10.1145/3453483.3454074}. Major benefits of this connection
in an MPC setting include better language abstractions for defining
protocols and for mechanisation and even automation of security proofs.
The goal of this paper is to develop a PL model for defining a variety of
low-level probabilistic MPC protocols, to formulate a collection
of confidentiality and integrity hyperproperties for our model
model with familiar information flow analogs, and to show how these
properties can be leveraged for improved proof automation.

The distinction between high- and low-level languages for MPC is
important. High-level languages such as Wysteria \cite{} and Viaduct
\cite{10.1145/3453483.3454074} are designed to provide effective
programming of full applications. These language designs incorporate
sophisticated verified compilation techniques such as orchestration
\cite{viaduct-UC} to guarantee high-level security properties, and
they rely on \emph{libraries} of low-level MPC protocols, such as
binary and arithmetic circuits. These low-level protocols encapsulate
abstractions such as secret sharing and semi-homomorphic encrypytion,
and rely on probabilistic progamming methods. So, low-level MPC
programming and protocol verification is a distinct challenge and both
critical to the general challenge of PL design for MPC and
complementary to high-level language design.

The connection between information flow hyperproperties and MPC
security is also complicated especially at a low level.  MPC protocols
involve communication between a group of distributed clients called a
\emph{federation} that collaboratively compute and publish the result
of some known \emph{ideal functionality} $\idealf$, maintaining
confidentiality of inputs to $\idealf$ without the use of a trusted
third party. However, since the outputs of $\idealf$ are public, some
information about inputs is inevitably leaked. Thus, the ideal
functionality establishes a declassification policy
\cite{sabelfeld2009declassification}, which is more difficult to
enforce than pure noninterference.  And subtleties of, e.g.,
semi-homomorphic encryption are central to both confidentiality and
integrity properties of protocols and similarly difficult to track
with coarse-grained security types alone.

Nevertheless, as previous authors have observed
\cite{5a51987acaa84c43bb4bf5bcc7d01683}, low-level protocol design
patterns such as secret sharing and circuit gate structure have
compositional properties that can be independently verified and then
leveraged in larger proof contexts. We contribute to this line of work
by developing an automated verification technique for subprotocols and
show how it can be integrated as a tactic in a larger security proof.

@@

As mentioned above, several high-level languages have been developed
for writing MPC applications, and frequently exploit the connection
between hyperproperties and MPC security. Previous work on analysis
for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
properties of complex MPC circuits, in particular a user-friendly
specification and automated enforcement of declassification bounds in
programs that use MPC in subprograms. The Wys$^\star$ language
\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
similar goals and includes a trace-based semantics for reasoning about
the interactions of MPC protocols. Their compiler also guarantees that
underlying multi-threaded protocols enforce the single-threaded source
language semantics. These two lines of work were focused on passive
security. The Viaduct language
\cite{10.1145/3453483.3454074} has a well-developed
information flow type system that automatically enforces both
confidentiality and integrity through hyperproperties such as robust
declassification, in addition to rigorous compilation guarantees
through orchestration \cite{viaduct-UC}. However, these high level
languages lack probabilistic features and other abstractions of
low-level protocols, the implementation and security of which are
typically assumed as a selection of library compnonents.

Various related low-level languages with probabilistic features have
also been developed. The $\lambda_{\mathrm{obliv}}$ language
\cite{darais2019language} uses a type system is used to automatically
enforce so-called probabilistic trace obliviousness.  But similar to
previous work on oblivious data structures \cite{10.1145/3498713},
obliviousness is related to pure noninterference, not the relaxed form
related to passive MPC security. The Haskell-based security type system in
\cite{6266151} enforces a version of conditioned noninterence that is
sound for passive security in a probabilisitic low-level setting, but
they do not consider malicious security. And properties of real/ideal
passive and malicious security for a probabilistic language have been
formulated in EasyCrypt \cite{8429300}-- though their proof methods,
while mechanized, are fully manual, and their formulation of malicious
security is not as clearly related to robust declassification as the
one we present in Section \ref{section-hyperprop}.

Program logics for probabilistic languages and specifically reasoning
about properties such as joint probabilistic independence is also
important related work. Probabilistic Separation Logic (PSL)
\cite{barthe2019probabilistic} develops a logical framework for
reasoning about probabilistic independence (aka separation) in
programs, and they consider several (hyper)properties, such as perfect
secrecy of one-time-pads and indistinguishability in secret sharing,
that are critical to MPC. However, their methods are manual, and
don't include conditional independence (separation). This
latter issue has been addressed in Lilac \cite{li2023lilac}. The
application of Lilac-style reasoning to MPC protocols has not
previously been explored, as we do in Section
\ref{section-example-gmw}.

@@

\paragraph{Fully and partially automated verification.} In Section
\ref{section-bruteforce} we develop a method for automatically
computing the pmf of $\minifed$ protocols in $\mathbb{F}_2$, that can
be automatically queried to enforce hyperperproperties of
security. This method is perfectly accurate but has high complexity;
we show this can be partially mitigated by conversion of protocols in
$\mathbb{F}_2$ to stratified Datalog which is amenable to HPC
acceleration. Furthermore, in Section \ref{section-examples-gmw} we
consider in detail how this automated technique can be used as a local
automated tactic for proving security in arbitrarily large GMW
circuits using conditional probabilistic independence as in
\cite{li2023lilac}.
